%%
%%

\section*{Installing and Configuring MySQL}
\label{_ChapterStart}
\index[general]{MySQL!Installing and Configuring }
\index[general]{Installing and Configuring MySQL }
\addcontentsline{toc}{section}{Installing and Configuring MySQL}

\subsection*{Installing and Configuring MySQL -- Phase I}
\index[general]{Installing and Configuring MySQL -- Phase I }
\index[general]{Phase I!Installing and Configuring MySQL -- }
\addcontentsline{toc}{subsection}{Installing and Configuring MySQL -- Phase I}

If you use the ./configure \verb{--{with-mysql=mysql-directory statement for
configuring {\bf Bacula}, you will need MySQL version 3.23.33 or later
installed in the {\bf mysql-directory} (we are currently using 3.23.56). If
MySQL is installed in the standard system location, you need only enter {\bf
\verb{--{with-mysql} since the configure program will search all the standard
locations. If you install MySQL in your home directory or some other
non-standard directory, you will need to provide the full path to it. 

Installing and Configuring MySQL is not difficult but can be confusing the
first time. As a consequence, below, we list the steps that we used to install
it on our machines. Please note that our configuration leaves MySQL without
any user passwords. This may be an undesirable situation if you have other
users on your system. 

Please note that as of Bacula version 1.31, the thread safe version of the
MySQL client library is used, and hence you must add the {\bf
\verb{--{enable-thread-safe-client} option to the {\bf ./configure} as shown below: 

\begin{enumerate}
\item Download MySQL source code from 
   \elink{www.mysql.com/downloads}{http://www.mysql.com/downloads}  

\item Detar it with something like:

   {\bf tar xvfz mysql-filename}  

Note, the above command requires GNU tar. If you do not  have GNU tar, a
command such as:

{\bf zcat mysql-filename | tar xvf - 

}  will probably accomplish the same thing. 

\item cd {\bf mysql-source-directory}

   where you replace {\bf mysql-source-directory} with the  directory name where
you put the MySQL source code.  

\item ./configure \verb{--{enable-thread-safe-client \verb{--{prefix=mysql-directory

   where you replace {\bf mysql-directory} with the directory  name where you
want to install mysql. Normally for system  wide use this is /usr/local/mysql.
In my case, I use  \~{}kern/mysql.  

\item make

   This takes a bit of time.  

\item make install

   This will put all the necessary binaries, libraries and support  files into
the {\bf mysql-directory} that you specified above.  

\item ./scripts/mysql\_install\_db

   This will create the necessary MySQL databases for controlling  user access.
Note, this script can also be found in the  {\bf bin} directory in the
installation directory 

\end{enumerate}

The MySQL client library {\bf mysqlclient} requires the gzip compression
library {\bf libz.a} or {\bf libz.so}. If you are using rpm packages, these
libraries are in the {\bf libz-devel} package. On Debian systems, you will
need to load the {\bf zlib1g-dev} package. If you are not using rpms or debs,
you will need to find the appropriate package for your system. 

At this point, you should return to completing the installation of {\bf
Bacula}. Later after Bacula is installed, come back to this chapter to
complete the installation. Please note, the installation files used in the
second phase of the MySQL installation are created during the Bacula
Installation. 
\label{mysql_phase2}

\subsection*{Installing and Configuring MySQL -- Phase II}
\index[general]{Installing and Configuring MySQL -- Phase II }
\index[general]{Phase II!Installing and Configuring MySQL -- }
\addcontentsline{toc}{subsection}{Installing and Configuring MySQL -- Phase
II}

At this point, you should have built and installed MySQL, or already have a
running MySQL, and you should have configured, built and installed {\bf
Bacula}. If not, please complete these items before proceeding. 

Please note that the {\bf ./configure} used to build {\bf Bacula} will need to
include {\bf \verb{--{with-mysql=mysql-directory}, where {\bf mysql-directory} is the
directory name that you specified on the ./configure command for configuring
MySQL. This is needed so that Bacula can find the necessary include headers
and library files for interfacing to MySQL. 

{\bf Bacula} will install scripts for manipulating the database (create,
delete, make tables etc) into the main installation directory. These files
will be of the form *\_bacula\_* (e.g. create\_bacula\_database). These files
are also available in the \lt{}bacula-src\gt{}/src/cats directory after
running ./configure. If you inspect create\_bacula\_database, you will see
that it calls create\_mysql\_database. The *\_bacula\_* files are provided for
convenience. It doesn't matter what database you have chosen;
create\_bacula\_database will always create your database. 

Now you will create the Bacula MySQL database and the tables that Bacula uses.


\begin{enumerate}
\item Start {\bf mysql}. You might want to use the {\bf startmysql}  script
   provided in the Bacula release.  
\item cd \lt{}install-directory\gt{}

   This directory contains the Bacula catalog  interface routines.  

\item ./grant\_mysql\_privileges

   This script creates unrestricted access rights for  {\bf kern}, {\bf kelvin},
and {\bf bacula}. You may  want to modify it to suit your situation. Please
note that  none of these userids including root are password protected.  

\item ./create\_mysql\_database

   This script creates the MySQL {\bf bacula} database.  The databases you create
as well as the access databases  will be located in \lt{}install-dir\gt{}/var/
in a subdirectory  with the name of the database, where \lt{}install-dir\gt{}
is the  directory name that you specified on the {\bf \verb{--{prefix} option. This 
can be important to know if you want to make a special backup  of the Bacula
database or to check its size.  

\item ./make\_mysql\_tables

   This script creates the MySQL tables used by {\bf Bacula}. 
\end{enumerate}

Each of the three scripts (grant\_mysql\_privileges, create\_mysql\_database
and make\_mysql\_tables) allows the addition of a command line argument. This
can be useful for specifying the user and or password. For example, you might
need to add {\bf -u root} to the command line to have sufficient privilege to
create the Bacula tables. 

To take a closer look at the access privileges that you have setup with the
above, you can do: 

\footnotesize
\begin{verbatim}
mysql-directory/bin/mysql -u root mysql
select * from user;
\end{verbatim}
\normalsize

\subsection*{Re-initializing the Catalog Database}
\index[general]{Database!Re-initializing the Catalog }
\index[general]{Re-initializing the Catalog Database }
\addcontentsline{toc}{subsection}{Re-initializing the Catalog Database}

After you have done some initial testing with {\bf Bacula}, you will probably
want to re-initialize the catalog database and throw away all the test Jobs
that you ran. To do so, you can do the following: 

\footnotesize
\begin{verbatim}
  cd <install-directory>
  ./drop_mysql_tables
  ./make_mysql_tables
\end{verbatim}
\normalsize

Please note that all information in the database will be lost and you will be
starting from scratch. If you have written on any Volumes, you must write and
end of file mark on the volume so that Bacula can reuse it. Do so with: 

\footnotesize
\begin{verbatim}
   (stop Bacula or unmount the drive)
   mt -f /dev/nst0 rewind
   mt -f /dev/nst0 weof
\end{verbatim}
\normalsize

Where you should replace {\bf /dev/nst0} with the appropriate tape drive
device name for your machine. 

\subsection*{Linking Bacula with MySQL}
\index[general]{Linking Bacula with MySQL }
\index[general]{MySQL!Linking Bacula with }
\addcontentsline{toc}{subsection}{Linking Bacula with MySQL}

After configuring Bacula with 

./configure \verb{--{enable-thread-safe-client \verb{--{prefix=\lt{}mysql-directory\gt{}
where \lt{}mysql-directory\gt{} is in my case {\bf /home/kern/mysql}, you may
have to configure the loader so that it can find the MySQL shared libraries.
If you have previously followed this procedure and later add the {\bf
\verb{--{enable-thread-safe-client} options, you will need to rerun the {\bf
ldconfig} program shown below. If you put MySQL in a standard place such as
{\bf /usr/lib} or {\bf /usr/local/lib} this will not be necessary, but in my
case it is. The description that follows is Linux specific. For other
operating systems, please consult your manuals on how to do the same thing: 

First edit: {\bf /etc/ld.so.conf} and add a new line to the end of the file
with the name of the mysql-directory. In my case, it is: 

/home/kern/mysql/lib/mysql then rebuild the loader's cache with: 

/sbin/ldconfig If you upgrade to a new version of {\bf MySQL}, the shared
library names will probably changes, and you must re-run the {\bf
/sbin/ldconfig} command so that the runtime loader can find them. 

Alternatively, your system my have a loader environment variable that can be
set. For example, on a Solaris system where I do not have root permission, I
use: 

LD\_LIBRARY\_PATH=/home/kern/mysql/lib/mysql 

Finally, if you have encryption enabled in MySQL, you may need to add {\bf
-lssl -lcrypto} to the link. In that case, you can either export the
appropriate LDFLAGS definition, or alternatively, you can include them
directly on the ./configure line as in: 

\footnotesize
\begin{verbatim}
LDFLAGS="-lssl -lcyrpto" \
   ./configure \
      <your-options>
\end{verbatim}
\normalsize
