%%
%%

\section*{System Requirements}
\label{_ChapterStart51}
\index[general]{System Requirements }
\index[general]{Requirements!System }
\addcontentsline{toc}{section}{System Requirements}

\label{SysReqs}

\subsection*{System Requirements}
\index[general]{System Requirements }
\index[general]{Requirements!System }
\addcontentsline{toc}{subsection}{System Requirements}

\begin{itemize}
\item {\bf Bacula} has been compiled and run on Linux RedHat, FreeBSD,  and
   Solaris systems. 
\item It requires GNU C++ version 2.95 or higher to compile. You can try  with
   other compilers and older versions, but you are on your own.  We have
   successfully compiled and used Bacula on RH8.0/RH9/RHEL 3.0  with GCC 3.2.
Note, in general GNU C++ is a separate package (e.g.  RPM) from GNU C, so you
need them both loaded. On RedHat systems,  the C++ compiler is part of the
{\bf gcc-c++} rpm package. 
\item There are certain third party packages that Bacula needs.  Except for
   MySQL and PostgreSQL, they can all be found in the  {\bf depkgs} and {\bf
   depkgs1} releases. 
\item If you want to build the Win32 binaries, you will need a  Microsoft
   Visual C++ compiler (or Visual Studio).  Although all components build
   (console has  some warnings), only the File daemon has been tested. 
\item {\bf Bacula} requires a good implementation of pthreads to work.  This
   is not the case on some of the BSD systems. 
\item The source code has been written with portability in mind and is  mostly
   POSIX compatible. Thus porting to any POSIX compatible  operating system
   should be relatively easy. 
\item The GNOME Console program is developed and tested under GNOME 2.x. It 
   also runs under GNOME 1.4 but this version is deprecated and  thus no longer
   maintained. 
\item The wxWidgets Console program is developed and tested with the latest 
   stable version of 
   \elink{wxWidgets}{http://www.wxwidgets.org/} (2.4.2).  It works fine with the
Windows and GTK+-1.x version of wxWidgets, and should  also works on other
platforms supported by wxWidgets. 
\item The Tray Monitor program is developed for GTK+-2.x. It needs  Gnome less
   or equal to 2.2, KDE greater or equal to 3.1 or any window manager supporting
   the  
\elink{ FreeDesktop system tray
standard}{http://www.freedesktop.org/Standards/systemtray-spec}. 
\item If you want to enable command line editing and history, you will  need
   to have /usr/include/termcap.h and either the termcap or the  ncurses library
   loaded (libtermcap-devel or ncurses-devel). 
\item If you want to use DVD as backup medium, you will need to download  and
   install the  
   \elink{dvd+rw-tools}{http://fy.chalmers.se/~appro/linux/DVD+RW/}. 
\end{itemize}
