%%
%%

\section*{Supported Operating Systems}
\label{_ChapterStart46}
\index[general]{Systems!Supported Operating }
\index[general]{Supported Operating Systems }
\addcontentsline{toc}{section}{Supported Operating Systems}

\subsection*{Supported Operating Systems}
\label{SupportedOSes}
\index[general]{Systems!Supported Operating }
\index[general]{Supported Operating Systems }
\addcontentsline{toc}{subsection}{Supported Operating Systems}

\begin{itemize}
\item Linux systems (built and tested on RedHat Enterprise Linux 3.0).  
\item If you have a recent Red Hat Linux system running the 2.4.x kernel  and
   you have the directory {\bf /lib/tls} installed on your  system (normally by
   default), bacula will {\bf NOT} run. This is  the new pthreads library and it
is defective. You must remove  this directory prior to running Bacula, or you
can simply change  the name to {\bf /lib/tls-broken}) then you must reboot
your  machine (one of the few times Linux must be rebooted). If  you are not
able to remove/rename /lib/tls, an alternative is to  set the environment
variable ``LD\_ASSUME\_KERNEL=2.4.19'' prior to  executing Bacula. For this
option, you do not need to reboot, and  all programs other than Bacula will
continue to use /lib/tls.  

The feedback that we have for 2.6 kernels is that the  same problem exists.
However, on 2.6 kernels, we would  probably recommend using the environment
variable override  (LD\_ASSUME\_KERNEL=2.4.19) rather than removing /lib/tls. 

\item Most flavors of Linux (Gentoo, SuSE, Mandrake, Debian, ...).  
\item Solaris various versions.  
\item FreeBSD (tape driver supported in 1.30 -- please see some  {\bf
   important} considerations in the 
   \ilink{ Tape Modes on FreeBSD}{FreeBSDTapes}  section of the
   Tape Testing chapter of this manual.)  
\item Windows (Win98/Me, WinNT/2K/XP) Client (File daemon) binaries.  
\item MacOS X/Darwin (see \elink{ http://fink.sourceforge.net/}{http://fink.sourceforge.net/} for
   obtaining the packages)  
\item OpenBSD Client (File daemon).  
\item Irix Client (File daemon).  
\item Tru64  
\item Bacula is said to work on other systems (AIX, BSDI, HPUX,  ...) but we
   do not have first hand knowledge  of these systems. 
\item  RHat 7.2 AS2, AS3, AS4, Fedora Core 2, SuSE SLES 7,8,9 and Debian Woody and Sarge Linux on
   S/390 and Linux on zSeries.
\item See the Porting chapter of the Bacula Developer's Guide  for information
   on porting to other systems. 
   \end{itemize}
