%%
%%

\section*{Thanks}
\label{_ChapterStart44}
\index[general]{Thanks }
\addcontentsline{toc}{section}{Thanks}

Thanks to Richard Stallman for starting the Free Software movement and for
bringing us gcc and all the other GNU tools as well as the GPL license. 

Thanks to Linus Torvalds for bring us Linux. 

Thanks to all the Free Software programmers. Without being able to peek at
your code, and in some cases, take parts of it, this project would have been
much more difficult. 

Thanks to John Walker for suggesting this project, giving it a name,
contributing software he has written, and for his programming efforts on
Bacula as well as having acted as a constant sounding board and source of
ideas. 

Thanks to the apcupsd project where I started my Free Software efforts, and
from which I was able to borrow some ideas and code that I had written. 

Special thanks to D. Scott Barninger for writing the bacula RPM spec file,
building all the RPM files and loading them onto Source Forge. This has been a
tremendous help.

Many thanks to Karl Cunningham for converting the manual from html format to
LaTeX. It was a major effort flawlessly done that will benefit the Bacula
users for many years to come. Thanks Karl.

Thanks to Dan Langille for the {\bf incredible} amount of testing he did on
FreeBSD. His perseverance is truly remarkable. Thanks also for the many
contributions he has made to improve Bacula (pthreads patch for FreeBSD,
improved start/stop script and addition of Bacula userid and group, stunnel,
...), his continuing support of Bacula users. He also wrote the PostgreSQL
driver for Bacula and has been a big help in correcting the SQL. 

Thanks to Phil Stracchino for writing the gnome-console {\bf ConsoleFont}
configuration command, all the suggestions he has made, and his continuing
suppport of Bacula users. 

Thanks to multiple other Bacula Packagers who make and release packages for
different platforms for Bacula. 

Thanks to Christopher Hull for developing the native Win32 Bacula emulation
code and for contributing it to the Bacula project. 

Thanks to Nicolas Boichat for writing wx-console and the bacula-tray-monitor.
These are very nice GUI additions to Bacula. 

Thanks to Nic Bellamy for providing the bacula-dir.conf file that he uses to
implement daily tape rotation using multiple Pools. 

Thanks to Johan Decock for providing numerous corrections to the manual. 

Thanks to all the Bacula users, especially those of you who have contributed
ideas, bug reports, patches, and new features. 

The original variable expansion code used in the LabelFormat comes from the
Open Source Software Project (www.ossp.org). It has been adapted and extended
for use in Bacula. 

For all those who I have left out, please send me a reminder, and in any case,
thanks for your contribution. 

\subsection*{Copyrights and Trademarks}
\index[general]{Trademarks!Copyrights and }
\index[general]{Copyrights and Trademarks }
\addcontentsline{toc}{subsection}{Copyrights and Trademarks}

Certain words and/or products are Copyrighted or Trademarked such as Windows
(by Microsoft). Since they are numerous, and we are not necessarily aware of
the details of each, we don't try to list them here. However, we acknowledge
all such Copyrights and Trademarks, and if any copyright or trademark holder
wishes a specific acknowledgment, notify us, and we will be happy to add it
where appropriate. 
